---
title: Cloud Computing
author: Jerzy Pawlowski
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: html_document
---


### Create a *Google Cloud* (gcloud) project

Create a virtual machine *Instance* on the *Google Compute Engine* (GCE).
https://cloud.google.com/compute/docs/quickstart-linux

Google cloud getting started community
https://cloud.google.com/community/
https://cloud.google.com/gcp/getting-started/

Google cloud console
https://console.cloud.google.com

Project dashboard info
https://console.cloud.google.com/home/dashboard?authuser=2&project=r-server-229012


<br>

### Configure the GCE virtual machine *Instance*

Google cloud console
https://console.cloud.google.com
https://console.cloud.google.com/compute/instances?project=r-server-229012


Enable OS Login - stores RSA keys for faster login
Open cloud console:
https://cloud.google.com/compute/docs/instances/managing-instance-access
Add in metadata:  enable-oslogin=TRUE

Open secure shell: in cloud console open ssh shell by selecting "Remote access" from drop-down menu next to instance
In Linux shell:
```{bash eval=FALSE}
whoami # algoquant_algoquants_ch
# Get Linux version and release
cat /etc/issue
cat /etc/os-release
uname -a
arch
# Set password
sudo passwd # star
```

Enable Compute Engine API - Create API Key Credentials for project r-server
https://console.cloud.google.com/apis/credentials?authuser=2&project=r-server-229012


<br>

### Log in remotely into the GCE virtual machine *Instance*

First method: Open secure ssh shell in cloud console by selecting "Remote access" from drop-down menu
Then in Linux shell:
```{bash eval=FALSE}
whoami # algoquant_algoquants_ch
# Get Linux version and release
cat /etc/issue
cat /etc/os-release
```

Second method (for faster login): First create secure key pair on Windows
https://www.ssh.com/ssh/keygen/
In Windows powershell (as Admin):
```{bash eval=FALSE}
# In powershell create secure key pair
ssh-keygen
```
Private key is in file C:\Users\Jerzy\.ssh\id_rsa
Public key is in file C:\Users\Jerzy\.ssh\id_rsa.pub
Change (remove) permissions for file id_rsa by right-clicking Properties

Copy public key to Linux account under username: algoquant_algoquants_ch
```{bash eval=FALSE}
# After OS-login is set up run gcloud command:
gcloud compute os-login ssh-keys add --key-file C:\Users\Jerzy\.ssh\id_rsa.pub
```
Or in ssh shell launched from cloud console:
```{bash eval=FALSE}
# Edit file authorized_keys and copy public key into it
sudo mkdir ~/.ssh
sudo nano ~/.ssh/authorized_keys
# Find the Linux server IP address
curl ifconfig.me
whoami # algoquant_algoquants_ch
```
Then in Windows powershell:
```{bash eval=FALSE}
# Run ssh to log into Linux
ssh algoquant_algoquants_ch@35.229.92.170
```


<br>

### Install software on the GCE virtual machine *Instance*

Install g++ on Linux
https://linuxconfig.org/how-to-install-g-the-c-compiler-on-ubuntu-18-04-bionic-beaver-linux
```{bash eval=FALSE}
sudo apt install build-essential
g++ --version
```

Install Java on Linux
```{bash eval=FALSE}
sudo apt-get install default-jre
sudo apt-get install default-jdk
sudo apt autoremove
java -version
```

Install libcurl on Linux
```{bash eval=FALSE}
# Using aptitude (better choice - allows resolving, configuring package dependencies)
sudo aptitude install -f libxml2-dev
sudo aptitude install -f libcurl4-openssl-dev
# Using apt-get
sudo apt-get install -f curl
sudo apt-get install -f libxml2-dev
sudo apt-get install -f libcurl4-openssl-dev
sudo apt-get update
sudo apt autoremove
```

Install OpenSSL on Linux
```{bash eval=FALSE}
# Using aptitude
sudo aptitude install -f libssl-dev
sudo aptitude install -f openssl-dev
# Using apt-get
sudo apt-get install -f libssl-dev
sudo apt-get update
sudo apt autoremove
```

Install Chromium browser on Linux
```{bash eval=FALSE}
# Using apt
sudo apt install -y chromium-browser
# Launch Chromium from ssh
DISPLAY=:1 sudo chromium-browser
```


<br>

### Install *R* on the GCE virtual machine *Instance*

Install R on Linux
https://cran.r-project.org/bin/linux/ubuntu/README.html
https://www.digitalocean.com/community/tutorials/how-to-install-r-on-ubuntu-18-04
```{bash eval=FALSE}
# download authentication keys from repository
sudo apt-key adv --keyserver keyserver.ubuntu.com --recv-keys E298A3A825C0D65DFD57CBB651716619E084DAB9
# add the repository
sudo add-apt-repository 'deb [arch=amd64,i386] https://cloud.r-project.org/bin/linux/ubuntu/cosmic-cran35/'
# run update to include package manifests from the new repository
sudo apt-get update

sudo rm /etc/apt/sources.list
sudo touch /etc/apt/sources.list

sudo nano /etc/apt/sources.list
# add
# deb https://cloud.r-project.org/bin/linux/ubuntu cosmic-cran35/

sudo apt-get update
sudo apt-get autoremove
# install base R
sudo apt-get install -f r-base
# install devel R
sudo apt-get install -f r-base-dev

# If above fails, this is alternative method using aptitude instead of apt-get
# https://askubuntu.com/questions/378558/unable-to-locate-package-while-trying-to-install-packages-with-apt/378565#378565
sudo add-apt-repository main
sudo add-apt-repository universe
sudo add-apt-repository restricted
sudo add-apt-repository multiverse
sudo apt-get update
sudo apt-get install aptitude
# install base R
sudo aptitude install -f r-base
# install devel R
sudo aptitude install -f r-base-dev

# update Linux packages
sudo apt-get update
sudo apt-get upgrade

```

Run R on Linux
```{bash eval=FALSE}
# run R as root
sudo R
```
Then install R packages
```{r eval=FALSE}
# in R install R packages
install.packages("RCurl")
install.packages("curl")
install.packages("git2r")
install.packages("openssl")
install.packages("httr")
install.packages("gh")
install.packages("usethis")
install.packages("devtools")
```

Then install more R packages
```{r eval=FALSE}
# in R install R packages
devtools::install_github("rstudio/httpuv")
install.packages(c("data.table", "zoo", "xts", "jsonlite", "dygraphs", "later", "promises", "httpuv", "mime", "rlang", "shiny", "htmltools", "knitr", "rmarkdown", "Rcpp", "RcppArmadillo", "matrixStats", "profvis", "ISLR", "DEoptim", "PerformanceAnalytics", "PortfolioAnalytics"), type="source")

devtools::install_github(repo="joshuaulrich/quantmod", type="source")
devtools::install_github(repo="joshuaulrich/TTR", type="source")

# Next install the packages rutils, HighFreq, and roll from GitHub as follows:
devtools::install_github(repo="algoquant/rutils", type="source")
devtools::install_github(repo="algoquant/HighFreq", type="source")
```


Install RStudio on Linux
https://www.rstudio.com/products/rstudio/download/
```{bash eval=FALSE}
# Install libapparmor1 and gdebi
sudo apt-get install libapparmor1 gdebi-core
# Download RStudio for 64-bit Ubuntu - check rstudio.com which is newest version
wget https://download1.rstudio.org/rstudio-xenial-1.1.463-amd64.deb -O rstudio.deb
# Install RStudio for 64-bit Ubuntu
sudo gdebi rstudio.deb
# Launch RStudio from ssh
DISPLAY=:1 sudo rstudio
```

Install RStudio server on Linux
https://www.rstudio.com/products/rstudio/download-server/
```{bash eval=FALSE}
# Download RStudio server for 64-bit Ubuntu - check rstudio.com which is newest version
wget https://download2.rstudio.org/rstudio-server-1.1.463-amd64.deb -O rstudio.deb
# Install RStudio for 64-bit Ubuntu
sudo gdebi rstudio.deb
rm rstudio.deb
```


<br>

### Install *VNC server* and *XFCE* on the GCE virtual machine *Instance*

Set firewall rules in Linux shell:
https://www.digitalocean.com/community/tutorials/initial-server-setup-with-ubuntu-18-04
```{bash eval=FALSE}
sudo ufw app list
# Allow ssh connections
sudo ufw allow OpenSSH
# Enable the firewall
sudo ufw enable
sudo ufw status
```

Note: VNC has problems running Chrome and RStudio
Install VNC server and XFCE in Linux shell:
https://www.digitalocean.com/community/tutorials/how-to-install-and-configure-vnc-on-ubuntu-18-04
```{bash eval=FALSE}
sudo apt update
# Install VNC server and XFCE
sudo apt install xfce4 xfce4-goodies
sudo apt install tightvncserver
# Start VNC server
vncserver
# Kill VNC server
vncserver -kill :1
# Reset VNC password
sudo vncpasswd
# Install autocutsel package to allow copy/paste
sudo apt-get install autocutsel
# List VNC startup script
cat ~/.vnc/xstartup
# Back up VNC startup script
sudo mv ~/.vnc/xstartup ~/.vnc/xstartup.bak
# Create new VNC startup script using nano editor:
sudo nano ~/.vnc/xstartup
# In nano add these lines without the "# "
# #!/bin/bash
# xrdb $HOME/.Xresources
# autocutsel -fork
# startxfce4 &
# Make VNC startup script executable
sudo chmod +x ~/.vnc/xstartup
# Start VNC server
vncserver
# Show open sockets ports
ss -ltn
# Kill VNC server (after VNC session is closed)
vncserver -kill :1
```

Install VNC viewer on Windows
https://www.realvnc.com/en/
In Windows powershell create an ssh tunnel by connecting local host to Linux:
```{bash eval=FALSE}
# Run ssh to log into Linux
ssh -L 5901:127.0.0.1:5901 -C -N -l algoquant_algoquants_ch 35.229.92.170
```
In VNC viewer app connect to local host (ignore encryption warning):
localhost:5901
A remote login window to Linux should open on Windows.


Configure Chromium browser on Linux: 
In VNC viewer window in Linux go to Applications menu on upper left -> Settings -> Preferred Applications -> select Chromium


<br>

### Install *X2Go* for remote desktop to Linux

Note: X2Go has problems running RStudio
Install X2Go server in Linux shell:
https://wiki.x2go.org/doku.php/doc:installation:x2goserver
```{bash eval=FALSE}
sudo add-apt-repository ppa:x2go/stable
sudo apt-get update
sudo apt-get install x2goserver x2goserver-xsession
```

Install X2Go client on Windows
https://wiki.x2go.org/doku.php/doc:installation:x2goclient
https://draculaservers.com/tutorials/install-x2go-ubuntu-remote-desktop/

Run X2Go client:
In Windows open X2Go client and create new sesssion (no need to tunnel ssh):
Host  35.229.92.170
Login algoquant_algoquants_ch
SSH port  22
RSA key C:/Users/Jerzy/.ssh/id_rsa
Session type XFCE

A remote login window to Linux should open on Windows.

