---
title: "Dashboard of Interactive Time Series Plots"
author: "Jerzy Pawlowski at NYU Tandon"
affiliation: NYU Tandon School of Engineering
email: jp3900@nyu.edu
date: '`r format(Sys.time(), "%m-%d-%Y %H:%M")`'
output:
  html_document:
    fig_width: 6
    fig_height: 4
---

```{r setup, include=FALSE}
# This is an R setup chunk, containing default options applied to all other chunks
knitr::opts_chunk$set(echo=FALSE)
library(rutils)
library(dygraphs)
```


### A dashboard of interactive time series plots written in *Rmarkdown*  

This *Rmarkdown* document contains interactive *dygraphs* plots of time series contained in the package *rutils*.  The interactive time series plots are produced using the package *dygraphs*.  It creates a list of *dygraphs* objects in a loop, and then uses the package *htmltools* to render (plot) the *dygraphs* objects.  

### You can click and drag your mouse to zoom into any plot.  

Double-click your mouse to restore the plot.


```{r echo=FALSE}
# Extract the closing ETF prices into a list of xts time series
# pricel <- lapply(rutils::etfenv$symbolv, function(symboln) {
#   quantmod::Cl(get(symboln, envir=rutils::etfenv))
# })  # end lapply

# Flatten (cbind) list of prices into single xts series
# prices <- rutils::do_call(cbind, pricel)

# Overwrite NA values
# prices <- zoo::na.locf(price)
# prices <- zoo::na.locf(prices, from_last=TRUE)

# Make nice column names
# colnames(prices) <- rutils::do_call_rbind(strsplit(colnames(prices), split="[.]"))[, 1]

# Create a list of dygraphs objects in loop
dyplots <- lapply(rutils::etfenv$symbolv, function(symboln) {
  pricev <- quantmod::Cl(get(symboln, envir=rutils::etfenv))
  dygraphs::dygraph(pricev, main=paste("Plot of:", rutils::get_name(colnames(pricev))), width=600, height=400)
})  # end lapply

# Load S&P500 stock prices
# load("/Users/jerzy/Develop/lecture_slides/data/sp500_prices.RData")
# Load old tickers
# sp500old <- read.csv(file="/Users/jerzy/Develop/lecture_slides/data/sp500_old.csv")
# symbolb <- sp500old$Ticker
# prico <- pricestock[, symbolb]
# Create a list of dygraphs objects in loop
# dyplots <- lapply(prico, function(pricev) {
#   dygraphs::dygraph(pricev, main=paste("Plot of:", colnames(pricev)), width=600, height=400)
# })  # end lapply

# Render the dygraphs objects
htmltools::tagList(dyplots)
```
