---
title: ioslides Presentations with Shiny
author: Jerzy Pawlowski
affiliation: NYU Tandon School of Engineering
date: '`r format(Sys.time(), "%m/%d/%Y")`'
email: jp3900@nyu.edu
css: slides.css
output: 
  ioslides_presentation: 
    widescreen: true
    smaller: true
    transition: 0.1
    self_contained: true
    logo: image/tandon_long_color.png
runtime: shiny
---

```{r setup, include=FALSE}
# This is an R setup chunk, containing default options applied to all other chunks
library(knitr)
# This sets the chunk default options
opts_chunk$set(cache=TRUE, collapse=TRUE, error=FALSE, prompt=TRUE)
# This sets the chunk display theme
knit_theme$set(knit_theme$get("acid"))
# This sets some display options
options(digits=3)
options(width=80)
```

[comment]: <> (This is a comment, it will not be included)
[//]: # (This is also a comment, and it may be the most platform independent comment)

<!-- regular html comment -->
<!--- special completely ignored comment -->


## Interactive *ioslides* Slide Presentations with Embedded *Shiny* Applications  

*Ioslides* is an interactive *html* slide presentation format developed by *RStudio*.  *Ioslides* presentations can have embedded *shiny* applications for interactive plots.  But slide presentations with embedded *shiny* interactive plots contain a live *R* session, so they can't be saved as a single file.  This presentation is an interactive *ioslides* presentation written in *R Markdown*, with embedded *shiny* applications written in *R* code.  

*RStudio* offers tutorials about writing presentations in *R Markdown*.
Introduction to *R Markdown*:
[*R Markdown*](http://rmarkdown.RStudio.com/)

Authoring *R Markdown* Presentations:
[Authoring *R Markdown* Presentations](https://support.RStudio.com/hc/en-us/articles/200486468-Authoring-R-Presentations)


*RStudio* offers several tutorials about writing interactive presentations in *R Markdown*.
Introduction to *ioslides*:
[*R Markdown*](http://rmarkdown.RStudio.com/ioslides_presentation_format.html)

This link offers a good introduction to *R Markdown* slide presentations, explaining the difference between *ioslides* presentations versus *slidy*:
[*R Markdown* slide presentation formats: *ioslides* versus *slidy*](http://data-analytics.net/cep/Schedule_files/presentations_demo.html)



## Interactive plots Using *Shiny*  

*Shiny* is an *R* package that allows creating interactive plots of *R* models.  A *shiny* visualization accepts user input of model parameters, recalculates the model, and displays an updated plot.  *Shiny* interactive plots contain a live *R* session, which allows them to recalculate an *R* model with user input model parameters. 

*RStudio* offers several tutorials about *shiny*.
*shiny* tutorial:
[*shiny* tutorial](http://shiny.RStudio.com/tutorial/)

Authoring *shiny* applications:
[authoring *shiny*](http://rmarkdown.RStudio.com/authoring_shiny.html)

Authoring embedded *shiny* applications:
[authoring embedded *shiny*](http://rmarkdown.RStudio.com/authoring_embedded_shiny.html)

Gallery of *shiny* applications:
[*shiny* gallery](http://shiny.RStudio.com/gallery/)

Hosting Interactive Presentations on shinyapps.io
[*shiny* apps](http://shiny.RStudio.com/articles/shinyapps.html)



## Slide with Embedded Inline *shiny* Application

Below is an embedded inline *shiny* application:

```{r, echo=FALSE, cache=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```



## Slide with Embedded External *Shiny* Application

Below is an embedded external *shiny* application, using function `shinyAppDir()` which executes *R* code contained in the files `ui.R` and `server.R`, stored in the normal_dist subdirectory:  

```{r, eval=TRUE, echo=TRUE, cache=FALSE}
library(shiny)
shinyAppDir(
  appDir="shiny/normal_dist",
  options=list(width="100%", height=400) # Doesn't change anything
  )  # end shinyAppDir
```


