---
title: Creating ioslides Presentations from R Markdown
author: Jerzy Pawlowski
affiliation: NYU Tandon School of Engineering
date: '`r format(Sys.time(), "%m/%d/%Y")`'
email: jp3900@nyu.edu
css: slides.css
output: 
  ioslides_presentation: 
    widescreen: true
    smaller: true
    transition: 0.1
    self_contained: true
    logo: image/tandon_long_color.png
---

```{r setup, include=FALSE}
# This is an R setup chunk, containing default options applied to all other chunks
library(knitr)
# This sets the chunk default options
opts_chunk$set(cache=TRUE, collapse=TRUE, error=FALSE, prompt=TRUE)
# This sets the chunk display theme - causes an error
# knit_theme$set(knit_theme$get("acid"))
# This sets some display options
options(digits=3)
options(width=80)
```

## *R Markdown* documents

This is an *ioslides* slide presentation rendered from an *R Markdown* document. Markdown is a simple markup language for publishing *html*, *pdf*, and MS Word documents. For more details on using *R Markdown* see <http://rmarkdown.rstudio.com>.


## Including *R* code chunks

*R* code chunks can be embedded in *R Markdown* documents, and they are automatically executed and their output is embedded in the output document.

```{r mtcars, echo=TRUE}
head(mtcars)
```


## Slide with interactive *dygraphs* plot

```{r, dygraphs_simple, cache=FALSE, eval=TRUE, echo=TRUE, fig.width=5, fig.height=4}
library(dygraphs)
dygraph(nhtemp, main = "New Haven Temperatures") %>% 
  dyRangeSelector(dateWindow = c("1920-01-01", "1960-01-01"))
```


## Two-column layout using HTML and CSS


<div class="column_left_40">
Example of two-column layout using `HTML` and `CSS`. 
The location of the break between columns is specified by the `CSS` style code in the file slides.css.
The left column is 40%, and the right is 60% of the page width. 

```{r, summ_cars, eval=FALSE, echo=TRUE, results="hold", size="tiny"}
summary(cars)
```

</div>

<div class="column_right_60">
Summary of `cars` data frame:

```{r, summ_cars, eval=TRUE, echo=FALSE, size="tiny"}
```

Note that the `echo = FALSE` parameter can be added to the code chunk to prevent printing of the *R* code that generated the plot.
</div>



## Plots in two-column layout using HTML and CSS

<div class="column_left_50">
Example of two-column layout with a plot.  
The Normal probability density function is given by:
$$
  P(x) = \frac{e^{-(x-\mu)^2/2\sigma^2}}{\sigma\sqrt{2 \pi}}
$$
Code for plotting the Normal probability density function:

```{r normal_dist_css, eval=FALSE, echo=(-(1:2)), results="hold"}
par(mar=c(2, 1, 2, 1))  # set plot margins
par(oma=c(1, 1, 1, 1))  # set outer margins
# plot a Normal probability distribution
curve(expr=dnorm(x, sd=1.5), type="l", xlim=c(-5, 5),
      xlab="", ylab="", lwd=2, col="blue", 
      main="Normal Density Function")
x_var <- seq(-5, 5, length=100)
y_var <- dnorm(x_var, mean=0.0, sd=1.5)
# set shading range
star_t <- 2; fin_ish <- 3
# set polygon base
are_a <- ((x_var >= star_t) & (x_var <= fin_ish))
# draw shaded area
polygon(c(star_t, x_var[are_a], fin_ish),
        c(-1, y_var[are_a], -1), col="red")
```

</div>

<div class="column_right_50">

Plot of Normal probability density function:

```{r, normal_dist_css, eval=TRUE, echo=FALSE, fig.width=5, fig.height=4}
```
Note that the `echo = FALSE` parameter can be added to the code chunk to prevent printing of the *R* code that generated the plot.
</div>



## Plots in two-column layout {.columns-2}

Example of two-column layout with a plot.  But there's no way to control the location of the content break between columns.

Plot the Normal probability density function:

```{r normal_dist, eval=TRUE, echo=TRUE, fig.width=5, fig.height=4}
par(mar=c(2, 1, 2, 1))  # set plot margins
par(oma=c(1, 1, 1, 1))  # set outer margins
x_var <- seq(-5, 7, length=100)
y_var <- dnorm(x_var, mean=1.0, sd=2.0)
plot(x_var, y_var, type="l", lty="solid",
     xlab="", ylab="")
title(main="Normal Density Function", line=0.5)
# set shading range
star_t <- 3; fin_ish <- 5
# set polygon base
are_a <- ((x_var >= star_t) & (x_var <= fin_ish))
# draw shaded area
polygon(c(star_t, x_var[are_a], fin_ish),
        c(-1, y_var[are_a], -1), col="red")
```

Note that the `echo = FALSE` parameter can be added to the code chunk to prevent printing of the *R* code that generated the plot.

## Two Columns - doesn't work {.smaller}  

<div class=".columns">

<div class=".column">
Column 1
</div>

<div class=".column">
Column 2
</div>

</div>

