---
title: Publishing R Markdown documents as pdf documents
author: algoquant
date: July 2, 2016
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```

### *R Markdown* documents

This is an *R Markdown* document. Markdown is a simple formatting syntax for publishing *HTML*, *pdf*, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.


### Including *R* code chunks

*R* code chunks can be embedded in *R Markdown* documents, and they are automatically executed and their output is embedded in the output document.

```{r mtcars, echo=TRUE}
head(mtcars)
```


### Including plots

You can also embed plots in *pdf* files, for example:

The Normal probability density function is given by:
$$
\displaystyle
  P(x) = \frac{e^{-(x-\mu)^2/2\sigma^2}}{\sigma\sqrt{2 \pi}}
$$

Plot the Normal probability density function:

```{r pressure, echo=TRUE}
x_var <- seq(-5, 7, length=100)
y_var <- dnorm(x_var, mean=1.0, sd=2.0)
plot(x_var, y_var, type="l", lty="solid",
     xlab="", ylab="")
title(main="Normal Density Function", line=0.5)
star_t <- 3; fin_ish <- 5  # set lower and upper bounds
# set polygon base
are_a <- ((x_var >= star_t) & (x_var <= fin_ish))
polygon(c(star_t, x_var[are_a], fin_ish),  # draw polygon
        c(-1, y_var[are_a], -1), col="red")
```

Note that the `echo = FALSE` parameter can be added to the code chunk to prevent printing of the *R* code that generated the plot.
