---
title: "Example R markdown Document"
author: Jerzy Pawlowski
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)

# install package quantmod if it can't be loaded successfully
if (!require("quantmod"))
  install.packages("quantmod")
```

### R markdown

This is an *R markdown* document. Markdown is a simple formatting syntax for authoring *HTML*, *pdf*, and *MS Word* documents. For more details on using *R markdown* see <http://rmarkdown.rstudio.com>.  

One of the advantages of writing documents *R markdown* is that they can be compiled into *HTML* documents, which can incorporate interactive plots,

You can read more about publishing documents using *R* here:  
https://algoquant.github.io/r,/markdown/2016/07/02/Publishing-documents-in-R/


You can read more about using *R* to create *HTML* documents with interactive plots here:  
https://algoquant.github.io/2016/07/05/Interactive-Plots-in-R/


You can read more markdown tutorials:  
http://daringfireball.net/projects/markdown/syntax  
https://guides.github.com/features/mastering-markdown/  
http://markdown-guide.readthedocs.io/en/latest/basics.html

Here are markdown cheatsheets:  
https://markdown-it.github.io/  
https://github.com/adam-p/markdown-here/wiki/Markdown-Cheatsheet  
https://daringfireball.net/projects/markdown/  
https://github.com/adam-p/markdown-here/wiki/Markdown-Cheatsheet  

This is a markdown interpreter:  
http://www.markitdown.net/markdown

You can read more about R markdown:  
http://rmarkdown.rstudio.com/
https://guides.github.com/features/mastering-markdown/


### The **knitr** package

The main difference between *markdown* and *R markdown*, is that *R markdown* allows including *R* code chunks.

Clicking the **knit** button in *RStudio*, compiles the *R markdown* document, including embedded *math formulas* and *R* code chunks, into the output document. 

Example of an *R* code chunk:

```{r cars}
summary(cars)
```


### Math formulas in *R markdown* documents

Math formulas can also be embedded in *R markdown* documents.  

For example inline formulas: $\frac{2}{3}$, $\sqrt{b^2 - 4ac}$, and $\hat{\lambda}=1.02$.  

Or display formulas (the Cauchy-Schwarz inequality):

$$
  \left( \sum_{k=1}^n a_k b_k \right)^2 
  \leq 
  \left( \sum_{k=1}^n a_k^2 \right) 
  \left( \sum_{k=1}^n b_k^2 \right) 
$$




### Plots in *R markdown* documents

Plots can also be embedded, for example:

```{r pressure, echo=TRUE}
plot(pressure)
```

Note that the `echo = FALSE` parameter may be added to the code chunk to prevent printing of the *R* code that generated the plot.




### Dynamic *dygraphs* plots in *R markdown* documents

*R markdown* allows including dynamic *dygraphs* plots if the output format is *HTML*. 
For example:

```{r dygraph, echo=TRUE}
dygraphs::dygraph(log(Cl(rutils::etfenv$VTI)), main="VTI Stock Price (log)")
```

