---
title: "Backtest of Contrarian SPY Strategy"
author_no_print: "Jerzy Pawlowski"
affiliation: NYU Tandon School of Engineering
date_no_print: '`r format(Sys.time(), "%m/%d/%Y")`'
email: jp3900@nyu.edu
output: html_document
runtime: shiny
---

This is a shiny app with a backtest of a contrarian trading strategy for 1-minute SPY returns.

The strategy calculates the z-scores from a rolling time series regression of 1-minute SPY returns against the time index, over a lookback period.

The rolling z-scores are calculated using the RcppArmadillo function HighFreq::roll_zscores().

The strategy identifies an oversold signal if the z-score is less than minus the threshold value, and an overbought signal if the z-score is greater than the threshold value.

The strategy trading rule is to buy (enter into a long position) if there is an oversold signal, and sell (enter a short position) if there is an overbought signal.

Then hold the long or short position until the next signal arrives.

The trading strategy has two parameters: the length of the lookback period and the threshold value.

The shiny app allows choosing the values of the lookback and the threshold parameters, and plots the cumulative PnL of the trading strategy.


```{r setup, include=FALSE}
# This is the R chunk setup code
knitr::opts_chunk$set(echo=TRUE)
library(HighFreq)  # load package HighFreq
# End R chunk setup code

# Data setup code
look_backs <- c(11, 111, 511, 1111, 5111)
price_s <- as.numeric(HighFreq::SPY[, "SPY.Close"])
re_turns <- rutils::diff_it(HighFreq::SPY[, "SPY.Close"])
load("C:/Develop/R/data/zscores_spy.RData")

# Define the strategy function
run_strategy <- function(re_turns, z_scores, thresh_old) {
  # Intitialize positions
  n_rows <- NROW(re_turns)
  position_s <- rep.int(NA_integer_, n_rows)
  position_s[1] <- 0
  # Handle signals
  bu_y <- (z_scores < (-thresh_old))
  position_s[bu_y] <- 1
  se_ll <- (z_scores > thresh_old)
  position_s[se_ll] <- (-1)
  position_s <- zoo::na.locf(position_s)
  # Lag the positions by two periods
  position_s <- rutils::lag_it(position_s, lagg=2)
  # Return strategy PnL
  cumsum(position_s*re_turns)
}  # end run_strategy


### Run the initial z-score data setup below once ###
# in_dex <- matrix(.index(HighFreq::SPY), nc=1)
# z_scores <- lapply(look_backs, function(look_back) {
#   z_score <- HighFreq::roll_zscores(res_ponse=re_turns, de_sign=in_dex, look_back=look_back)
#   HighFreq::roll_scale(z_score, look_back=look_back, use_median=TRUE)
# })  # end lapply
# z_scores <- rutils::do_call(cbind, z_scores)
# colnames(z_scores) <- paste0("look_back=", look_backs)
# z_scores[which(!is.finite(z_scores), arr.ind=TRUE)] <- 0
# save(z_scores, file="C:/Develop/R/data/zscores_spy.RData")
### End initial data setup ###
```


```{r trading_strategy, echo=FALSE}
# R chunk with shiny code

# Define the input panel
inputPanel(
  selectInput("look_back", label="lookback value:",
              choices=look_backs, selected=look_backs[4]),
  # actionButton("re_calculate", "Recalculate z-scores"),
  sliderInput("thresh_old", label="threshold value:",
              min=0.01, max=5.0, value=0.5, step=0.01)
)  # end inputPanel

cat("Please Wait While Model is Running...\n")

# Run the trading strategy and plot it
renderPlot({
  # Extract from input the strategy model parameters
  look_back <- paste0("look_back=", input$look_back)
  z_scores <- subset(z_scores, select=look_back)
  # thresh_old <- input$thresh_old*sum(abs(range(z_scores)))/2
  thresh_old <- input$thresh_old

  # Run the trading strategy and plot it
  pnl_s <- run_strategy(re_turns, z_scores, thresh_old)[endpoints(HighFreq::SPY, on="days")]
  chart_Series(pnl_s, name="Cumulative PnL of the Contrarian Strategy for 1-minute SPY Returns")

})  # end renderPlot
```
