---
title: "Backtest of Stock Momentum Strategy S&P500 Stocks"
author_no_print: "Jerzy Pawlowski"
affiliation: NYU Tandon School of Engineering
date_no_print: '`r format(Sys.time(), "%m/%d/%Y")`'
email: jp3900@nyu.edu
output: html_document
runtime: shiny
---

This is a shiny app with a backtest of a momentum trading strategy for daily S&P500 stock returns.

The strategy performs a rolling calculation of the portfolio weights based on the relative performance over a lookback period.

The rolling portfolio weights are calculated using the RcppArmadillo function HighFreq::back_test().

The strategy trading rule is to buy (enter into long positions) the stocks with the best relative performance, and sell (enter short positions) the worst.

Then hold the long or short positions until the next rebalancing end point (months, weeks, etc.)

The trading strategy has several parameters which are explained in the HighFreq package manual for the RcppArmadillo function HighFreq::calc_weights().

The shiny app allows choosing the values of the parameters, and plots the cumulative PnL of the trading strategy.


```{r setup, include=FALSE}
# This is the R chunk setup code
knitr::opts_chunk$set(echo=TRUE)
library(HighFreq)  # load package HighFreq
# End R chunk setup code

# Data setup code
load("/Users/jerzy/Develop/lecture_slides/data/sp500_returns.RData")
# load("/Users/jerzy/Develop/lecture_slides/data/sp500_prices.RData")
# Subset the columns with non-zero returns
returns <- returns[, !(returns[(NROW(returns) %/% 10), ] == 0)]
# Subset 100 columns to reduce computations
set.seed(1121)  # reset random number generator
sam_ple <- sample(1:NCOL(returns), 100)
returns <- returns[, sam_ple]
# excess <- returns_scaled[, sam_ple]
excess <- returns
stock_symbols <- colnames(returns)
ncols <- NCOL(returns)
endpoints <- rutils::calc_endpoints(returns, interval="weeks")
endpoints <- endpoints[endpoints > (ncols+1)]
nrows <- NROW(endpoints)

# Define the strategy function
run_strategy <- function(excess, returns, typev, look_back, alpha, max_eigen, quantilev, coeff, bid_offer, lagg) {
  # browser()
# cat("look_back =", look_back, "\nalpha =", alpha, "\nmax_eigen =", max_eigen, "\nlagg =", lagg, "\n")
  startpoints <- c(rep_len(1, look_back-1), endpoints[1:(nrows-look_back+1)])
  # Perform backtest in RcppArmadillo
  pnls <- HighFreq::back_test(excess=excess, 
                               returns=returns,
                               startpoints=startpoints-1,
                               endpoints=endpoints-1,
                               typev=typev,
                               alpha=alpha,
                               max_eigen=max_eigen,
                               quantilev=quantilev,
                               coeff=coeff,
                               bid_offer=bid_offer)
  
  pnls <- xts(cumsum(pnls), order.by=index(returns))
  prices <- cumsum(rowMeans(returns))
  pnls <- cbind(pnls, prices)
  colnames(pnls) <- c("Strategy", "Index")
  pnls
}  # end run_strategy


### Run the initial data setup below once ###
# load("/Users/jerzy/Develop/lecture_slides/data/sp500.RData")
# returns <- eapply(env_sp500, function(xt_s) {
#   rutils::diffit(log(quantmod::Cl(xt_s)))
# })  # end eapply
# returns <- rutils::do_call(cbind, returns)
# returns <- zoo::na.fill(returns, 0)
# sum(!is.finite(returns))
# colnames(returns) <- rutils::get_name(colnames(returns))
# save(returns, file="C:/Develop/R/data/returns_percent_sp500.RData")
### End initial data setup ###
```


```{r trading_strategy, echo=FALSE}
# R chunk with shiny code

# Define the input panel
inputPanel(
  selectInput("typev", label="Portfolio weights type:",
              choices=c("max_sharpe", "min_var", "min_varpca", "rank", "rankrob"), 
              selected="rank"),
  sliderInput("look_back", label="lookback (weeks):",
              min=2, max=75, value=55, step=1),
  # actionButton("re_calculate", "Recalculate z-scores"),
  sliderInput("alpha", label="Shrinkage intensity:",
              min=0, max=1, value=0.8, step=0.01),
  sliderInput("max_eigen", label="max eigenvalues:",
              min=2, max=20, value=15, step=1),
  sliderInput("quantilev", label="quantile:",
              min=0, max=0.5, value=0.1, step=0.01),
  numericInput("bid_offer", label="bid-offer:",
               value=0.001),
  selectInput("coeff", label="coeff:",
              choices=c(-1, 1), selected=1),
  sliderInput("lagg", label="lag:",
               min=1, max=10, value=1, step=1),
  actionButton("re_calculate", "Recalculate the Model")
)  # end inputPanel

# chart_Series(xts(cumsum(rowMeans(returns)), index(returns)),
#              name="Average Stock Prices")

pnls <- shiny::reactive({
  # Extract from input the strategy model parameters
  typev <- isolate(input$typev)
  look_back <- isolate(input$look_back)
  alpha <- isolate(input$alpha)
  max_eigen <- isolate(input$max_eigen)
  quantilev <- isolate(input$quantilev)
  bid_offer <- isolate(input$bid_offer)
  coeff <- isolate(as.numeric(input$coeff))
  lagg <- isolate(input$lagg)
  # Model is re-calculated when the re_calculate variable is updated
  input$re_calculate
  
  # Run the trading strategy
  cat("Please Wait While Model is Running...\n")
  pnls <- run_strategy(excess, returns, typev, look_back, alpha, max_eigen, quantilev, coeff, bid_offer, lagg)
  cat("Model finished.\n")
  pnls
})  # end reactive code


# Plot the trading strategy
renderDygraph({
    colnamevs <- colnames(pnls())
    captiont <- paste("Momentum Strategy for S&P500 Stocks")
    dygraphs::dygraph(pnls(), main=captiont) %>%
      dyAxis("y", label=colnamevs[1], independentTicks=TRUE) %>%
      dyAxis("y2", label=colnamevs[2], independentTicks=TRUE) %>%
      dySeries(name=colnamevs[1], axis="y", label=colnamevs[1], strokeWidth=1, col="red") %>%
      dySeries(name=colnamevs[2], axis="y2", label=colnamevs[2], strokeWidth=1, col="blue")
  })  # end output plot


# renderPlot({
#   chart_Series(pnls(), name="Cumulative PnL of a Stock Momentum Strategy")
#   # plot(cumsum(rnorm(1e2)), main="Cumulative PnL of a Stock Momentum Strategy")
# })  # end renderPlot

```
