---
title: "Backtest of Stock Momentum Strategy"
author_no_print: "Jerzy Pawlowski"
affiliation: NYU Tandon School of Engineering
date_no_print: '`r format(Sys.time(), "%m/%d/%Y")`'
email: jp3900@nyu.edu
output: html_document
runtime: shiny
---

This is a shiny app with a backtest of a momentum trading strategy for daily S&P500 stock returns.

The strategy performs a rolling calculation of the portfolio weights based on the relative performance over a lookback period.

The rolling portfolio weights are calculated using the RcppArmadillo function HighFreq::back_test().

The strategy trading rule is to buy (enter into long positions) the stocks with the best relative performance, and sell (enter short positions) the worst.

Then hold the long or short positions until the next rebalancing end point (months, weeks, etc.)

The trading strategy has several parameters which are explained in the HighFreq package manual for the RcppArmadillo function HighFreq::calc_weights().

The shiny app allows choosing the values of the parameters, and plots the cumulative PnL of the trading strategy.


```{r setup, include=FALSE}
# This is the R chunk setup code
knitr::opts_chunk$set(echo=TRUE)
library(HighFreq)  # load package HighFreq
# End R chunk setup code

# Data setup code
load("/Users/jerzy/Develop/R/data/returns_percent_sp500.RData")
# Subset 100 columns to reduce computations
set.seed(1121)  # reset random number generator
returns <- returns[, sample(1:NCOL(returns), 100)]
stock_symbols <- colnames(returns)
ncols <- NCOL(returns)
endp <- rutils::calc_endpoints(returns, interval="weeks")
endp <- endp[endp > (ncols+1)]
nrows <- NROW(endp)

# Define the strategy function
run_strategy <- function(returns, typev, look_back, alpha, dimax, quantilev, coeff, bid_offer, lagg) {
  # browser()
# cat("look_back =", look_back, "\nalpha =", alpha, "\ndimax =", dimax, "\nlagg =", lagg, "\n")
  startp <- c(rep_len(1, look_back-1), endp[1:(nrows-look_back+1)])
  # Perform backtest in RcppArmadillo
  pnls <- HighFreq::back_test(excess=returns, 
                               returns=returns,
                               startp=startp-1,
                               endp=endp-1,
                               typev=typev,
                               alpha=alpha,
                               dimax=dimax,
                               quantilev=quantilev,
                               coeff=coeff,
                               bid_offer=bid_offer)
  xts(cumsum(pnls), order.by=index(returns))
}  # end run_strategy


### Run the initial data setup below once ###
# load("/Users/jerzy/Develop/lecture_slides/data/sp500.RData")
# returns <- eapply(env_sp500, function(xtes) {
#   rutils::diffit(log(quantmod::Cl(xtes)))
# })  # end eapply
# returns <- rutils::do_call(cbind, returns)
# returns <- zoo::na.fill(returns, 0)
# sum(!is.finite(returns))
# colnames(returns) <- rutils::get_name(colnames(returns))
# save(returns, file="/Users/jerzy/Develop/R/data/returns_percent_sp500.RData")
### End initial data setup ###
```


```{r trading_strategy, echo=FALSE}
# R chunk with shiny code

# Define the input panel
inputPanel(
  selectInput("typev", label="Portfolio weights type:",
              choices=c("max_sharpe", "min_var", "min_varpca", "rank", "rankrob"), 
              selected="max_sharpe"),
  sliderInput("look_back", label="lookback (weeks):",
              min=2, max=75, value=55, step=1),
  # actionButton("recalcb", "Recalculate z-scores"),
  sliderInput("alpha", label="Shrinkage intensity:",
              min=0, max=1, value=0.8, step=0.01),
  sliderInput("dimax", label="max eigenvalues:",
              min=2, max=20, value=15, step=1),
  sliderInput("quantilev", label="quantile:",
              min=0, max=0.5, value=0.1, step=0.01),
  numericInput("bid_offer", label="bid-offer:",
               value=0.001),
  selectInput("coeff", label="coeff:",
              choices=c(-1, 1), selected=1),
  sliderInput("lagg", label="lag:",
               min=2, max=10, value=2, step=1)
)  # end inputPanel

cat("Please Wait While Model is Running...\n")
# chart_Series(xts(cumsum(rowMeans(returns)), index(returns)),
#              name="Average Stock Prices")
  
# Run the trading strategy and plot it
renderPlot({
  # Extract from input the strategy model parameters
  typev <- input$typev
  look_back <- input$look_back
  alpha <- input$alpha
  dimax <- input$dimax
  quantilev <- input$quantilev
  bid_offer <- input$bid_offer
  coeff <- as.numeric(input$coeff)
  lagg <- input$lagg

  # Run the trading strategy and plot it
  pnls <- run_strategy(returns, typev, look_back, alpha, dimax, quantilev, coeff, bid_offer, lagg)
  chart_Series(pnls, name="Cumulative PnL of a Stock Momentum Strategy")
  # plot(cumsum(rnorm(1e2)), main="Cumulative PnL of a Stock Momentum Strategy")

})  # end renderPlot
```
