---
title: Interactive Plotly Plots for Interest Rates and Yield Curve
author: Jerzy Pawlowski, NYU Tandon School of Engineering
date:   '`r format(Sys.time(), "%m/%d/%Y")`'
output: html_document
---

### Interactive *plotly* Treasury rates plot

The script (not displayed) loads time series of constant maturity Treasury rates, contained in an environment called *env_rates*.  
It then produces an interactive time series plot of the 20yr and 10yr Treasury rates, using the *plotly* package.

To select a date range, left-click on the plot and drag your mouse.
Then double-click (or click on home icon) to return to the original range.

```{r eval=TRUE, echo=FALSE, fig.width=7, fig.height=4}
# load quantmod and plotly
suppressMessages(suppressWarnings(library(quantmod)))
suppressMessages(suppressWarnings(library(plotly)))
# load constant maturity Treasury rates
load(file="C:/Develop/data/rates_data.RData")
# create data frame of time series
rates <- na.omit(do.call(merge, as.list(env_rates)[c("DGS10", "DGS20")]))
rates <- data.frame(dates=index(rates), coredata(rates))
# plotly syntax using pipes
rates %>% 
  plot_ly(x=~dates, y=~DGS10, type="scatter", mode="lines + markers", name="10yr Treasury rate") %>% 
  add_trace(x=~dates, y=~DGS20, name="20yr Treasury rate") %>% 
  layout(title="20yr and 10yr Treasury rates", 
         xaxis=list(title="Time"),
         yaxis=list(title="Rates"),
         legend=list(x=0.1, y=0.9))
# standard plotly syntax - for reference
# p_lot <- plot_ly(data=rates, x=~dates, y=~DGS10, type="scatter", mode="lines", name="10yr Treasury rate")
# p_lot <- add_trace(p=p_lot, x=~dates, y=~DGS20, fill="tonexty", name="20yr Treasury rate")
# p_lot <- layout(p=p_lot, title="20yr and 10yr Treasury rates", xaxis=list(title="Time"), yaxis=list(title="Rates"), legend=list(x=0.1, y=0.9))
# p_lot
```


### Please Ignore all below this line: Interactive *plotly* yield curve plot

The script loads constant maturity Treasury rates time series, contained in an environment called *env_rates*.  
It then produces a *plotly* interactive time series plot of the Treasury yield curve, to illustrate its changing shape over time.
using the *plotly* package.

```{r eval=FALSE, echo=FALSE, fig.width=7, fig.height=4}
# this script doesn't work

# load quantmod and plotly
suppressMessages(suppressWarnings(library(quantmod)))
suppressMessages(suppressWarnings(library(plotly)))

# load constant maturity Treasury rates
load(file="C:/Develop/data/rates_data.RData")
# get end-of-year dates since 2006
dates <- endpoints(env_rates$DGS1["2006/"], on="years")
dates <- index(env_rates$DGS1["2006/"])[dates]
# create time series of end-of-year rates
rates <- eapply(env_rates, function(ra_te) ra_te[dates])
rates <- do.call(merge, rates)
# rename columns and rows, sort columns, and transpose into matrix
colnames(rates) <- substr(colnames(rates), start=4, stop=11)
rates <- rates[, order(as.numeric(colnames(rates)))]
colnames(rates) <- paste0(colnames(rates), "yr")
rates <- t(rates)
colnames(rates) <- substr(colnames(rates), start=1, stop=4)
# plot matrix
color_ramp <- colorRampPalette(c("red", "blue"))(NCOL(rates))

# plotly syntax using pipes

# create yield_curves list
yield_curves <- lapply(1:NCOL(rates), function (colnum) {
  list(visible=FALSE,
       name=paste0('v=', colnamevs[colnum]),
       x=row_names,
       y=rates[, colnum])
})  # end lapply

yield_curves[3][[1]]$visible=TRUE

# create slider list
slide_r <- lapply(1:NCOL(rates), function (colnum) {
  ste_p <- list(args=list("visible", rep(FALSE, length(yield_curves))), method="restyle")
  ste_p$args[[2]][colnum] <- TRUE
  ste_p
})  # end lapply

# plot all traces
p_lot <- plot_ly()
for (colnum in 1:NCOL(rates)) {
  p_lot <- add_lines(p=p_lot, x=yield_curves[colnum][[1]]$x, 
                     y=yield_curves[colnum][[1]]$y, 
                     visible=yield_curves[colnum][[1]]$visible, 
                     name=yield_curves[colnum][[1]]$name, 
                     type="scatter", mode="lines", hoverinfo="name", 
                     line=list(color="00CED1"), showlegend=FALSE)

}  

# add layout with slider control
p_lot <- layout(p=p_lot, 
                title="20yr and 10yr Treasury rates", 
                xaxis=list(title="Time", 
                           rangeslider=list(list(active=3,
                                                 currentvalue=list(prefix="Frequency: "),
                                                 steps=slide_r))), 
                yaxis=list(title="Rates"), 
                legend=list(x=0.1, y=0.9))

p_lot

# standard plotly syntax - for reference
# p_lot <- plot_ly(data=rates, x=~dates, y=~DGS10, type="scatter", mode="lines", name="10yr Treasury rate")
# p_lot <- add_trace(p=p_lot, x=~dates, y=~DGS20, fill="tonexty", name="20yr Treasury rate")
# p_lot <- layout(p=p_lot, title="20yr and 10yr Treasury rates", xaxis=list(title="Time"), yaxis=list(title="Rates"), legend=list(x=0.1, y=0.9))
# p_lot
```

